Vragenlijst thesis evaluatie

### 2015-04-26 doc/template/chap-8.tex
aanpassing van hoofdstuk

\todo{resources vs nodes}
\section{Introduction}
This section gives a formal definition of the FlipIt game with a virus propagation. First we derive a formula for a FlipIt game without a virus. After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation. 
%In this section we are going to elaborate how we are going to model a Flipit game with multiple resources and a virus that propagates and infects the resources. We come up with a formula for the normal FlipIt (normal as in specific parameters and no normalising over the first interval) and then reform it to a FlipIt game with a virus.

\subsection{FlipIt vs FlipIt with virus propagation}
Section [] explained the FlipIt game. This section will expands the FlipIt game to a FlipIt game with virus propagation. A FlipIt game with virus propagation is a game with multiple nodes where the attacker will drop a virus on one of the nodes. The virus will then spread itself to the neighbour nodes trying to infect every node. The attacker will only gain control over the whole network -or the game in general- when it has infected all the nodes.

After dropping a virus on the first resource, it takes a while for the virus to infect the entire network. The time that it takes for the virus to infect every resource will be denoted as parameter \textit{d}. If we want to measure how long it takes for the virus to infect all the resources, we have to calculate the shortest path from the first infected node to the farthest node. This can be measured by a method that we will explain in section []. Using this parameter, the gain of a FlipIt game will be calculated in section []. \\ \todo{uitleggen waarom we een gewone flippit kunnen nemen met 1 resource}
\todo{a virus propagation}
The gain of a player is defined as the total amount of time that a player has owned the resource since the start of the game. In the context of a FlipIt game with virus propagation, the whole network is seen as one resource. This resource is owned by the attacked if he has control over the entire network. If the defender owns at least one node, he owns the resource.\\

@@

\todo{deze sectie herschrijven}
There is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt, but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control. \\
\todo{we hebben een periodisch spel, simpelste spel.}
\todo{ergens vermelden waarom we deze case bespreken en niet waarbij de aanvaller niet periodisch speelt}

\todo{zeggen dat we eerst de functie voorstellen van de normale FlipIt}
First a list of notations that will be used throughout the formal definition (see figure \ref{fig:notations} for a graphic representation of some of the notations):
\begin{description}
\item $\delta_{0}$: This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. 
\item $\delta_{1}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker.
\item \textit{$T_{0}$}: This denotes the phase of the defender that was random and uniform chosen over the interval [0,$\delta_{0}$].
\item \textit{$T_{1}$}: This denotes the phase of the attacker that was chosen uniformly at random in interval [0,$\delta_{1}$].
\item \textit{Unit of control}: Defined as the period between gaining control and losing control over the resource.  
\item \textit{n}: n is the n'th interval of the attacker, starting from interval 1.
\item $\Delta A$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker.
\item \textit{lcm(a,b)}: The \textit{lcm} of a and b is the least common multiple of a and b.
\item \textit{gcd(a,b)}: The \textit{gcd} of a and b is the greatest common divider of a and b.
\end{description}
\begin{figure}[hbtp]
\caption{Defining unit of control}
\centering
\includegraphics[scale=1]{Images/FlipSpel.png}
\label{fig:notations}
\end{figure}

@@

\todo{hier staat al benefit dat nog niet is uitgelegd}
We \todo{We mag gebruikt worden in een wetenschappelijke tekst zolang de focus blijft op het werk en niet op de schrijver, FlipIt schrijvers gebruiken ook veel de we vorm} start by computing the gain of the attacker in a periodic game without phases. After that we introduce the phases. Next we compute the benefit of the attacker in function of the gain. Finally we adapt our gain and benefit calculation to include virus propagation. 
\\
\subsubsection{Computing the gain for an attacker of a normal FlipIt game}
\textbf{Case  1:} \\
For $\delta_{1} > \delta_{0}$ (The defender moves faster than the attacker.) \\

@@

For that reason we divide our time line of our FlipIt game into different intervals of size $\delta_{1}$. Therefore every time the attacker moves we have the start of a new interval. Considering that the defender will move faster than the attacker, he or she will at least move one time during the interval of the attacker. Because the attacker only moves at the start of his or her interval we can say that the defender will always end as being in control of the resource. \todo{tekening met voorbeeld}. To calculate how long the unit of control of the attacker is in every interval, we only need to know how long the defender had control during that interval and substract it from the length of the interval of the attacker. 
\todo{blabla en dan hebben we de volgende formule}
This brings us to the next forumula to calculate the length of a unit of control in the n'th interval of the attacker. 
\todo{definieer dat delta groter als nul is en niet negatief}
For every real $\delta_{1}$ and $\delta_{0}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
\begin{equation}\label{first}
\Delta A = [( 1- n  ) \cdot \delta_{1}] mod \delta_{0}
\end{equation}
where n is the number of the n'th interval of the attacker where the length of the unit of control of the attacker is calculated.\\

@@

\textbf{Case 2}:
Alles omdraaien \todo{beter beschrijven}
\subsubsection{Computing the average benefit for a normal FlipIt game}

The average benefit of a player \textit{i} is denoted by $\beta_{i}(t) = \gamma_{i}(t) - k_{i} / \delta_{i}$, which is equal to the fraction of time that the resource has been owned by player \textit{i}, minus the cost rate for moving. For now we consider the cost rate equal to 0. $\gamma_{i}(t) = G_{i}(t)/t$ is the average gain rate of player \textit{i} which is defined as the fraction of time that player \textit{i} has control over the resource up to time t.
As stated before the formula that will be computed is in function of $\delta_{1}$. So parameter t will be in this case a multiple of the period of the attacker, hence a multiple of $\delta_{1}$. In the rest of the paper the average benefit will be addressed as the benefit of the game for player \textit{i}. \\

@@

\subsection{Formula with a virus propagation}
Now we can define how we can use the previous formula to calculate the benefit of the attacker with a virus propagation.
As mentioned before parameter \textit{d} defines the virus propagation. It will take an amount of time \textit{d} before the attacker gains control over all the resources. In the previous section we defined a formula to calculate each unit of control of the attacker. If the virus propagation takes \textit{d} time before every resource is infected then this d has to be subtracted from each unit of control. (see figure \ref{fig:virusflip}). It may happen that the unit of control is less than \textit{d}. The result of the substration will be a negative number in time. In this case this means that the defender has flipped all the resources before the attacker could gain control over all the resources. To calculate the benefit only the units of control bigger than 0 have to be summarized. So the formula becomes:

\begin{equation}\label{first}
Gain_{1} = \dfrac{\sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace }{delta_{0} \cdot delta_{1}} 
\end{equation}

>>>>

\todo{resources vs nodes}
\section{Introduction}

%This section gives a formal definition of the FlipIt game with a virus propagation. First we derive a formula for a FlipIt game without a virus. After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation. 
In this section we are going to elaborate how we are going to model a FlipIt game with multiple resources and a virus that propagates and infects the resources. First we derive a formula for a FlipIt game without a virus. After that we introduce a modification to this formula to achieve an adapted formula for a FlipIt game with a virus propagation. 
% We come up with a formula for the normal FlipIt (normal as in specific parameters and no normalising over the first interval) and then reform it to a FlipIt game with a virus.

\subsection{FlipIt vs FlipIt with virus propagation}
Section [] explained the FlipIt game. This section will adapt a FlipIt game to a FlipIt game with virus propagation. We consider the non-adaptive continuous FlipIt game where both players play a periodic strategy with a random phase. De adaptation will be a game where the single resource consist of a computer network with multiple nodes. One of the players, the defender, will try to defend his network. The defender will do this by flipping all the nodes of the network in every move he plays. The attacker, the other player, will try to infect all the nodes in the network. The attacker will do this by flipping the node in the graph that can infect all the nodes in the shortest time possible. The attacker will only gain the control over the resource when he infects all the nodes of the network. The defender in contrast will gain control over the whole network when it has control over at least one node in the network. \\
\todo{ergens vermelden waarom we deze case bespreken en niet waarbij de aanvaller niet periodisch speelt}

After dropping a virus on the first resource, it takes a while for the virus to infect the entire network. The time that it takes for the virus to infect every node will be denoted as parameter \textit{d}. If we want to measure how long it takes for the virus to infect all the nodes in the network, we have to calculate the shortest path from the first infected node to the farthest node. This can be measured by a method that we will explain in section []. Using this parameter, a FlipIt game with virus propagation can be modelled. \\

@@

\todo{deze sectie herschrijven}
There is a definition given for the gain of a player \textit{i} by the writers of the paper FlipIt [], but we want to add the property of a virus propagation to the game, hence parameter \textit{d}, so we are trying to find another formula that defines a game by counting the amount of time one of the players has control. \\
\todo{we hebben een periodisch spel, simpelste spel.}

\todo{zeggen dat we eerst de functie voorstellen van de normale FlipIt}
First a list of notations that will be used throughout the formal definition (see figure \ref{fig:notations} for a graphic representation of some of the notations):
\begin{description}
\item $\delta_{0}$: Positive and non zero number. This is the period of the defender. This denotes the length of the interval between two consecutive moves of the defender. 
\item $\delta_{1}$: This is the period of the attacker. This denotes the length of the interval between two consecutive moves of the attacker.
\item \textit{$T_{0}$}: This denotes the phase of the defender that was random and uniform chosen over the interval [0,$\delta_{0}$].
\item \textit{$T_{1}$}: This denotes the phase of the attacker that was chosen uniformly at random in interval [0,$\delta_{1}$].
\item \textit{Unit of control}: Defined as the period between gaining control and losing control over the resource.  
\item \textit{n}: n is the n'th interval of the attacker, starting from interval 1.
\item $\Delta A$ : This is a function that denotes the length of a unit of control of the attacker in the n'th interval of the attacker.
\item \textit{d}: Amount of time that the virus needs to infect every node in the network.
\item \textit{lcm(a,b)}: The \textit{lcm} of a and b is the least common multiple of a and b.
\item \textit{gcd(a,b)}: The \textit{gcd} of a and b is the greatest common divider of a and b.
\item \textit{Gain}:  The gain of a player is defined as the total amount of time that a player has owned the resource since the start of the game. In the context of a FlipIt game with virus propagation, the whole network is seen as one resource. This resource is owned by the attacked if he has control over the entire network. If the defender owns at least one node, he owns the resource.\\
\item \textit{Average Gain}:  $\gamma_{i}(t) = G_{i}(t)/t$ is the average gain rate of player \textit{i} which is defined as the fraction of time that player \textit{i} has control over the resource up to time t.
\item \textit{Benefit}: The average benefit of a player \textit{i} is denoted by $\beta_{i}(t) = \gamma_{i}(t) - k_{i} / \delta_{i}$, which is equal to the fraction of time that the resource has been owned by player \textit{i}, minus the cost rate for moving. For now we consider the cost rate equal to 0. In the rest of the paper the average benefit will be addressed as the benefit of the game for player \textit{i}.
\end{description}
\begin{figure}[hbtp]
\caption{Defining unit of control}
\centering
\includegraphics[scale=1]{Images/FlipSpel.png}
\label{fig:notations}
\end{figure}

@@

\todo{We mag gebruikt worden in een wetenschappelijke tekst zolang de focus blijft op het werk en niet op de schrijver, FlipIt schrijvers gebruiken ook veel de we vorm} 
We start by computing the gain of the attacker in a periodic game without phases. Two cases are considered: case 1 where the defender moves as least as fast as the attacker and case 2 where the attacker plays as least as fast as the defender. After that we introduce the phases. Next we adapt our gain calculation to include virus propagation. Finally we compute the benefit of the attacker in function of the gain.   
\\
\subsubsection{Computing the gain for an attacker of a normal FlipIt game}
\textbf{Case  1:} \\
For $\delta_{1} > \delta_{0}$ (The defender moves faster than the attacker.) \\

@@

For that reason we divide our time line of our FlipIt game into different intervals of size $\delta_{1}$. Therefore every time the attacker moves we have the start of a new interval. Considering that the defender will move faster than the attacker, he or she will at least move one time during the interval of the attacker. Because the attacker only moves at the start of his or her interval we can say that the defender will always end as being in control of the resource. \todo{tekening met voorbeeld}. To calculate how long the unit of control of the attacker is in the n'th interval, we only need to know how long the attacker has control over the resource before the defender moves in that interval. First we need to calculate the start time of the n'th interval which will be a multiple of the period of the attacker. Once we know this we need to know what the difference is between this amount of time and the amount of time when the defender will flip. To do this we take the modulo of the negative of the amount of time of the attacker with the period of the defender. By doing so we calculate how much it takes to add to this time to compleet to a time that is a multiplier of the period of the defender. See figure \ref{fig:modulo} to see a graphic representation of this modulo. 
\todo{blabla en dan hebben we de volgende formule}
\begin{figure}[hbtp]
\caption{Taking the modulo of a negative number}
\centering
\includegraphics[scale=0.6]{Images/Modulo.png}
\label{fig:modulo}
\end{figure}

\todo{modulo uitleggen en dan verder door tot deze formule}
This brings us to the next forumula to calculate the length of a unit of control in the n'th interval of the attacker. 
\todo{definieer dat delta groter als nul is en niet negatief}
For every positive and non zero real $\delta_{1}$ and $\delta_{0}$ $\in$ \(\mathbb{R}\) and every n $\in$ \(\mathbb{N}\) (including 0 in the set of natural numbers) :
\begin{equation}\label{first}
\Delta A = [( 1- n  ) \cdot \delta_{1}] mod \delta_{0}
\end{equation}
where n is the number of the n'th interval of the attacker where the length of the unit of control of the attacker is calculated.\\

@@

\textbf{Case 2}:
Alles omdraaien \todo{beter beschrijven}

\subsection{Formula with a virus propagation}
Now we can define how we can use the previous formula to calculate the gain of the attacker in a FlipIt game with virus propagation.
As mentioned before parameter \textit{d} defines the virus propagation. It will take an amount of time \textit{d} before the attacker gains control over all the resources. In the previous section we defined a formula to calculate each unit of control of the attacker. If the virus propagation takes \textit{d} time before every resource is infected then this d has to be subtracted from each unit of control. (see figure \ref{fig:virusflip}). It may happen that the unit of control is less than \textit{d}. The result of the substration will be a negative number in time. In this case this means that the defender has flipped all the resources before the attacker could gain control over all the resources. To calculate the gain only the units of control bigger than 0 have to be summarized. So the formula becomes:

\begin{equation}\label{first}
Gain_{1} = \sum_{i=0}^{\delta_{0}} \lbrace [( 1 - i ) \cdot \delta_{1}] mod \delta_{0} - d \rbrace  > 0 \rbrace 
\end{equation}

\subsubsection{Computing the benefit of a FLipIt game with virus propagation}

@@
